---
title: "Homework 3"
author: "[YOUR NAME HERE]"
date: "`r Sys.Date()`"
output: 
  html_document:
    code_folding: hide
    df_print: paged
    fig_caption: yes
    fig_height: 6
    fig_width: 8
    highlight: textmate
    theme: readable
    toc: yes
    toc_float: yes
bibliography: bibliography.bib
---

<!-- Feel free to edit theme / highlight in the YAML header above -->

```{r setup, echo=TRUE, message = FALSE, warning = FALSE}

knitr::opts_chunk$set(echo = TRUE)

# Load libraries for homework problems
library(tidyverse)
library(gt)

# Read in the data
abpm_wide <- read_csv('data/abpm_wide_synthetic.csv')

```


# Overview

Ambulatory blood pressure monitoring (ABPM) is a technique for assessing a person’s blood pressure (BP). ABPM is conducted with a special device that consists of a BP cuff worn on the participant's arm and attached to a small recording device worn on the belt. The ABPM device is usually worn for 24 hours, and it records BP periodically (usually at 15-minute or 30-minute intervals). One of the benefits of ABPM is measuring BP during routine daily activities and during sleep instead of in clinical settings. A previous study  by @hermida2018asleep found that, among all BP derive risk factors, asleep Systolic BP is the most associated with cardiovacular disease events. 

## Data dictionary

I have modified the New York times data to include information about state's population levels. The data are described below: 

```{r}

c("sleep_time" = "Time of falling asleep", 
  "awake_time" =  "Time of waking up", 
  "age" = "Participant age, years",
  "sex" = "Participant sex at birth",
  "race" = "Participant race",
  "educ" = "Participant education at exam time",
  "smoke" = "Participant smoking status at exam time",
  "sbp_0 - sbp_23" = "Systolic BP, hours since midnight",
  "dbp_0 - dbp_23" = "Diastolic BP, hours since midnight",
  "hr_0 - hr_23" = "Heart rate, hours since midnight"
) %>% 
  enframe() %>% 
  gt(rowname_col = "name") %>%
  tab_stubhead(label = 'Variable name') %>% 
  cols_label(value = 'Variable description') %>% 
  cols_align('right') %>%
  tab_source_note("BP = blood pressure") %>% 
  tab_header(title = 'Dictionary for synthetic ABPM data')
  
```

## Data

The data (`abpm_wide`) are printed below:

```{r}

abpm_wide

```

# Problem 1

Convert the `smoke` variable in `abpm_wide` into a factor and exclude participants with missing data for `sleep_time`, `awake_time`, and `smoke`.

__Notes__: 

The factor labels should be

- Never smoked
- Former smoker
- Current smoker

```{r}

read_rds('solutions/01_solution.rds')

```

# Problem 2

Pivot the data into a longer format.

__Notes__:

- Create a column named `id` that uniquely identifies each row of `abpm_wide`.

- Pivot the data into a longer form such that each `id` has a column for `sbp`, `dbp`, and `hr`. You may need to use `pivot_longer`, then `separate`, then `pivot_wider`.

- Drop all rows with missing data for `sbp`, `dbp`, _or_ `hr`

```{r}

read_rds('solutions/02_solution.rds')

```


# Problem 3

Create a factor variable called `awake` that has values of `'Yes'` when participants are awake and `'No'` when asleep.

__Notes__: 

- There are two scenarios that are relevant: 

    + sleep time is less than awake time
    
    + sleep time is greater than awake time.
    
- Be sure to think about how the `awake` indicator should be defined in _both_ of these scenarios.

- The `if_else()` function is helpful here.

```{r, fig.width=10}

read_rds('solutions/03_solution.rds')

```


# Problem 4

Exclude participants with less than 10 ABPM readings while awake _or_ less than 5 ABPM readings while asleep.

__Notes__: 

- You may do this with or without using `pivot` functions. 

```{r}

read_rds('solutions/04_solution.rds')

```

# Problem 5

Using the data from problem 4, compute the mean systolic BP, diastolic BP, and heart rate for each participant during their awake and asleep periods, separately.

__Notes__: 

- After summarizing your data, you will want to use `pivot_wider`. 

- Before using `pivot_wider`, I'd recommend recoding the `awake` indicator variable so that values of `'Yes'` are mapped to `'awk'` and values of `'No'` are mapped to `'slp'`. 

- Remember to `ungroup()` your data if necessary.

```{r}

read_rds('solutions/05_solution.rds')

```

# Problem 6

Tabulate __and interpret__ your findings. Show the mean systolic BP, diastolic BP, and heart rate while awake and asleep for males and females, stratified by race and smoking status, jointly. (see my table for an illustration of joint stratification). 

__Notes__: 

- To get full credit for this exercise, write _three_ observations about your results. Here is an example (you cannot use this as one of your own observations): 

    + BP during sleep tends to be lower than BP during wakefulness. 

- `unite()` is a helpful function for joint stratification.  

- You can rename your columns and use `tab_spanner_delim` for convenience.

```{r}

read_rds('solutions/06_solution.rds')

```

# References




