---
title: "Homework 3"
author: "[YOUR NAME HERE]"
date: "`r Sys.Date()`"
output: 
  html_document:
    code_folding: hide
    df_print: paged
    fig_caption: yes
    fig_height: 6
    fig_width: 8
    highlight: textmate
    theme: readable
    toc: yes
    toc_float: yes
bibliography: bibliography.bib
---

<!-- Feel free to edit theme / highlight in the YAML header above -->

```{r setup, echo=TRUE, message = FALSE, warning = FALSE}

knitr::opts_chunk$set(echo = TRUE)

# Load libraries for homework problems
library(tidyverse)
library(gt)

# Read in the data
abpm_wide <- read_csv('data/abpm_wide_synthetic.csv')

```


# Overview

Ambulatory blood pressure monitoring (ABPM) is a technique for assessing a person’s blood pressure (BP). ABPM is conducted with a special device that consists of a BP cuff worn on the participant's arm and attached to a small recording device worn on the belt. The ABPM device is usually worn for 24 hours, and it records BP periodically (usually at 15-minute or 30-minute intervals). One of the benefits of ABPM is measuring BP during routine daily activities and during sleep instead of in clinical settings. A previous study  by @hermida2018asleep found that, among all BP derive risk factors, asleep Systolic BP is the most associated with cardiovacular disease events. 

## Data dictionary

I have modified the New York times data to include information about state's population levels. The data are described below: 

```{r}

c("sleep_time" = "Time of falling asleep", 
  "awake_time" =  "Time of waking up", 
  "age" = "Participant age, years",
  "sex" = "Participant sex at birth",
  "race" = "Participant race",
  "educ" = "Participant education at exam time",
  "smoke" = "Participant smoking status at exam time",
  "sbp_0 - sbp_23" = "Systolic BP, hours since midnight",
  "dbp_0 - dbp_23" = "Diastolic BP, hours since midnight",
  "hr_0 - hr_23" = "Heart rate, hours since midnight"
) %>% 
  enframe() %>% 
  gt(rowname_col = "name") %>%
  tab_stubhead(label = 'Variable name') %>% 
  cols_label(value = 'Variable description') %>% 
  cols_align('right') %>%
  tab_source_note("BP = blood pressure") %>% 
  tab_header(title = 'Dictionary for synthetic ABPM data')
  
```

## Data

The data (`cv19`) are printed below:

```{r}

abpm_wide

```

# Problem 1

Convert the `smoke` variable in `abpm_wide` into a factor and exclude participants with missing data for `sleep_time`, `awake_time`, and `smoke`.

__Notes__: 

The factor labels should be

- Never smoked
- Former smoker
- Current smoker

```{r}

read_rds('solutions/01_solution.rds')

```

# Problem 2

Pivot the data into a longer format.

__Notes__:

- Create a column named `id` that uniquely identifies each row of `abpm_wide`.

- Pivot the data into a longer form such that each `id` has a column for `sbp`, `dbp`, and `hr`. You may need to use `pivot_longer`, then `separate`, then `pivot_wider`.

- Drop all rows with missing data for `sbp`, `dbp`, _or_ `hr`

```{r}

read_rds('solutions/02_solution.rds')

```


# Problem 3

Create an factor variable called `awake` that has values of `'Yes'` when participants are awake and `'No'` when asleep.

__Notes__: 

- There are two scenarios that are relevant: 

    + sleep time is less than awake time
    
    + sleep time is greater than awake time.

```{r, fig.width=10}

read_rds('solutions/03_solution.rds')

```


# Problem 4

__Notes__: 

```{r}

read_rds('solutions/04_solution.rds')

```

# Problem 5

__Notes__: 

```{r}

read_rds('solutions/05_solution.rds')

```





